Dynastic Extinctions and Urban Economic Development: \\Evidence from the Holy Roman Empire\footnote{I thank Davide Cantoni, Matthias Weigand, and my RA colleagues at the Chair of Economic History for helpful comments and feedback.}

### 2023-06-08 paper/paper.tex
#22 explain key assumption

Past treatments of this topic have suffered from a lack of available historical data. \cite{diamond1997} and \cite{landes2006} each present narrative arguments for the economic benefits of state competition but stop short of formal empirical analysis. Such a study requires at the very least three things: 1) a detailed territorial history of European cities, 2) some city-level measure of development reaching far enough into the past, 3) some measure of state power, institutional quality, or whatever else one wishes to explain effects with.
These data limitations are starting to be overcome. Recently \cite{schoenholzer2022} have investigated how switching between states affects the development of cities. They find that switching states affects population negatively in the short run but positively in the long run. Similar to the evolution metaphor, they interpret this as "creative destruction": the short run cost of destruction that is compensated by the long-run benefit of becoming part of a "better" state.
However, they do not directly estimate the effect of anything other than switching to different states.

I do this by looking at the development of cities after they switch states. I take their analysis a step further by identifying special cases of switches: in \textit{conquests} a city is forcefully absorbed into a new state, while in \textit{successions} a city is absorbed into a new state after the rulers of the old state have gone extinct. Comparing the effects of these two types of switches allows me to tease out the effect of state power.

I make a key assumption: conquests imply state power differentials but successions do not. If state A takes a city from state B, 
absorbing a city by conquest is more likely if the new state is more powerful than the old one, but absorbing a city by succession underlies no such limitations - it only requires 
the new state after a conquest is selected for state power, but the new state after a succession is not. Roughly speaking, 
If a city is absorbed via conquest,

The dataset I use has several advantages. First, it traces the territorial history of cities in the Holy Roman Empire precisely up to a single year. Second, it contains

Rather than m

@@

\section{Historical Background}

>>>>

Past treatments of this topic have suffered from a lack of available historical data. \cite{diamond1997} and \cite{landes2006} each present narrative arguments for the economic benefits of state competition but stop short of formal empirical analysis. Such a study requires at the very least three things: 1) a detailed territorial history of European cities, 2) some city-level measure of development reaching far enough into the past, 3) some measure of state power, institutional quality, or whatever else one wishes to explain effects with.
These data limitations are starting to be overcome. Recently \cite{schoenholzer2022} have investigated how switching between states affects the development of cities. They find that switching states affects population negatively in the short run but positively in the long run. Similar to the evolution metaphor, they interpret this as "creative destruction": the short run cost of destruction is compensated by the long-run benefit of becoming part of a "better" state. However, their

The analysis in this paper extends that of \cite{schoenholzer2022} by identifying special cases of switching events: in \textit{conquests} a city is forcefully absorbed into a new state, while in \textit{successions} a city is absorbed into a new state after the rulers of the old state have gone extinct. Comparing the effects of these two types of switches allows me to tease out the effect of state power.

I make a key assumption: conquests imply state power differentials but successions do not. In other words, if we consider the universe of all switching events, the average increase in state power from the old to the new owner would be higher for the conquests than for the successions. The reasoning is that conflict --- as in a conquest --- selects the new state for power, but conflict-free successions, depending mainly on dynastic marriages, are more "random". This assumption allows me to use conquests as a stand-in for an increase in state power.

The rest of the paper proceeds as follows.

@@

\section{Background}
